Propositions as Sessions, Mechanically

### 2015-02-25 paper/paper.tex
Fixed issue with bib/hyperref (I think). More thoughts on CP.

Another aspect of the encoding is how to define terms. The intrinsic encoding
described by~\citeauthor{Benton:2012:STT}~\cite{Benton:2012:STT} is one
approach, indexing terms by their type so as to prevent ill-formed terms from
being constructed. However, in the encoding for GV there are issues with using
this encoding: $(1)$ it is not immediately clear how to handle linear contexts
using the de Bruijn variable encoding, since in the intrinsic setting the
environment supports weakening; $(2)$ an intrinsic encoding of terms would be
complicated by the need to enforce that session types occur in certain
instances (by use of a predicate); and $(3)$ well-typed terms require extra
assumptions about binder freshness which cannot reasonably be expressed as a
function type. For these reasons, an intrinsic approach does not offer much
benefit since a well-typed term relation would still need to be
defined. Therefore, I chose an extrinsic encoding for terms and define
well-typed terms as a separate inductive type.

\citeauthor{Aydemir:2008:EFM}~\cite{Aydemir:2008:EFM} note that the size of
their language proof infrastructure is proportional to the number of binding
constructs. In the case of a simply typed lambda calculus this is not onerous,
but GV has four binding constructs, CP has six and propositions have two. It
may be true that this approach results in too many auxiliary definitions and
operations not relevant to the actual properties one wishes to prove. However,
I believe that once this infrastructure has been defined, the system can be
extended without concern for these binding constructs.

@@

Formalising the process calculus CP provided many challenges for the
metatheory library since to my knowledge, the library has never been applied
to a process calculus before. So for instance, in the lambda calculus one
typically does not work up to permutation of binders, whereas in a process
calculus these kinds of structural equivalences are standard. Additionally,
substitution and opening are only defined on processes for names. In other
words, one cannot substitute an arbitrary process term for a name, unlike in
the lambda calculus. Thus frameworks which assume a lambda calculus style
substitution operation are not immediately amenable to the CP
setting~\cite{Lee:2012}.

@@

Explicit weakening and contraction were required to differentiate the
reduction rules for servers and clients. In contrast, \citeauthor{Wadler:2014}
defines these operations explicitly.

@@

for an unbound index so this decision is sound. A consequence is that I must
define a notion of opening language terms involving binders as described
by~\citeauthor{Aydemir:2008:EFM}. For example, an abstraction with body term
$t$ which is well-typed if $t$ is well-typed when its unbound index,
representing the argument to the abstraction, is replaced by some fresh
variable $x$, written $t^x$. Some terms in GV require multiple binders but I
believe it will be straightforward to extend this notion. Unfortunately, I
must define similar operations on CP terms and handle substitution and opening
on them as well as propositions which can accept propositional variables. I
hope to be able to generalise as much as possible with the help of the
Metatheory library. say something about gmeta here and its automatic
generation of these infrastructure lemmas

@@

%\printbibliography

>>>>

Another aspect of the encoding is how to define terms. The intrinsic encoding
described by~\citeauthor{Benton:2012:STT}~\cite{Benton:2012:STT} is one
approach, indexing terms by their type so as to prevent ill-formed terms from
being constructed. However, in the encoding for CP/GV there are issues with
using this encoding: $(1)$ it is not immediately clear how to handle linear
contexts using the de Bruijn variable encoding, since in the intrinsic setting
the environment automatically supports weakening; $(2)$ an intrinsic encoding
of GV terms would be complicated by the need to enforce that session types
occur in certain instances (by use of a predicate); and $(3)$ well-typed terms
require extra assumptions about binder freshness which cannot reasonably be
expressed as a function type. For these reasons, an intrinsic approach does
not offer much benefit since a well-typed term relation would still need to be
defined. Therefore, I chose an extrinsic encoding for terms and define
well-typed terms as a separate inductive type.

\citeauthor{Aydemir:2008:EFM}~\cite{Aydemir:2008:EFM} note that the size of
their language proof infrastructure is proportional to the number of binding
constructs. In the case of a simply typed lambda calculus this is not onerous,
but GV has four binding constructs, CP has six and propositions have
two. Indeed, a lot of effort was expended setting up this
infrastructure. However, I believe one of the contributions of this work is to
articulate what one needs from state of the art metatheory libraries in order
to handle session-based languages, and more generally, linear type systems.

@@

Formalising the process calculus CP provided many challenges for the
metatheory library since to my knowledge, the library has never been applied
to a process calculus before. So for instance, in the lambda calculus one
typically does not work up to permutation of binders, whereas in a process
calculus these kinds of structural equivalences are standard. Additionally,
substitution and opening are only defined on processes for names. In other
words, one cannot substitute an arbitrary process term for a name, unlike in
the lambda calculus. Thus frameworks which assume a lambda calculi style
substitution operation are not immediately amenable to the CP
setting~\cite{Lee:2012}.

@@

Initially, an implicit notion of weakening and contraction were chosen so as
to be in line with \citeauthor{Wadler:2014}'s definitions. However, explicit
weakening and contraction were required to differentiate the reduction rules
for the various interactions between servers and clients.

Ordering in environments is ignored and the development required lemmas
allowing the permutation of environments. Support for permutating environments
is not currently supported in the Metatheory library but some general lemmas
were added as a result of this work. Additionally, custom tactics were
developed to automatatically solve some simple permutation goals. However,
these tactics are not a ``catch all'' and some generalising still needs to be
done.

The paper presentation omits details: the principal cut rules do not specify
weakening could occur: since weakening is implicit is tacitly assumed
weakening can always be pushed further up the derivation tree e.g. AxCut rule

another interesting aspect which the formalism reveal is the rule for
associativity is constructed in a way to permit associativity in a certain way
pre-determined

metatheory library and the library of tactics from software foundations are
used extensively: metatheory does not handle multiple binders (check this?)
maybe add a generic tactic to handle multiple binders... or perhaps add a
lemmas for handling binders

@@

@@

\sloppy
\printbibliography
